Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2018-01-05 paper/aldebaran.tex
precision

From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do precision asteroseismology from the ground and obtain very precise estimates of stellar parameters. Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis. Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements. One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g. ][]{frink2001,hekker2006,ortiz2016}. The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars. Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, at a sampling we find to be insufficient for our purposes, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case. This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.

>>>>

From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do asteroseismology from the ground and obtain very precise estimates of stellar parameters. Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis. Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements. One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g. ][]{frink2001,hekker2006,ortiz2016}. The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars. Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, at a sampling we find to be insufficient for our purposes, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case. This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.
